Interlocks Mitigate Information Cascades in Large-Scale Social Learning

### 2024-10-30 manuscript.tex
Updates from Overleaf

%\author{Herminio Bodon}
%\orcid{0009-0004-3270-3621}
%\email{HerminioBodon2020@u.northwestern.edu}
%\affiliation{%
%  \institution{Northwestern University}
%  \city{Evanston}
%  \state{Illinois}
%  \country{USA}
}
%\affiliation{%
%  \institution{University of Michigan}
%  \city{Ann Arbor}
%  \state{Michigan}
%  \country{USA}
}

@@

In this paper, we begin with a review of the literature on large-scale participatory decision-making and related topics.
We focus, in particular, on empirical studies of successful large-scale participatory projects, looking for commonalities.
We then focus on one such commonality: social network structure,
specifically interlock networks.
We also narrow our focus to social learning:
the process by which individuals revise their beliefs and opinions
based on observations and interactions with others---a core part of participatory decision-making.
Finally, we develop an agent-based model
to test the effects of interlock networks on social learning
in large groups.
Our findings suggest that interlock network structure may contribute to the success of large-scale participatory projects by providing protection from information cascades and polarization.

@@

\subsection{Collaboration and Network Structure}
As we will elaborate below,
we find that real-world examples of successful large-scale participatory collaborations
often share a common trait:
interlocking social network structure.
Here, we use ``network'' in the sense of a set of entities with dyadic relationships between them \cite{newman_networks_2018}, e.g.,
individuals linked to other individuals through direct communication.
Similarly, we use ``structure'' to refer to the configuration of individual dyadic relationships and connections rather than in the ``social system'' sense, e.g., the difference in communication links between a roundtable discussion and a lecture.
An {\em interlocking network} or {\em interlock}
is a network in which groups of entities are connected (interlocked)
through shared individual constituents \cite{taylor_world_2015}.
As a much-studied example,
an interlocking directorate is a network of corporations
in which two corporations (the groups) are considered connected if any individual director (the constituent) belongs to both of their boards 
\cite{mizruchi_what_1996}.
Or as a more modern example,
collaborative interactions on Wikipedia take place on individual articles, their per-article talk pages, and on topical WikiProject pages \cite{platt_network_2018}.
In this example, articles/talk pages/projects are the groups, and individual editors are the constituents.

@@

There is considerable evidence that group composition
\cite{hill_group_1982,
nishi_inequality_2015,
robert_jr_differences_2018,
sydow_diversity_2017,
lerner_diverse_2018,
arazy_information_2011}
and social network structure
\cite{
kearns_experiments_2012,
barkoczi_social_2016,
salehi_hive:_2018,
mason_propagation_2008,
mason_collaborative_2012}
can influence the success of a collaboration.
Interlocking structure has implications for both of these factors.
By dividing an entire collaborative project into smaller groups,
interlocks influence group size and group composition,
potentially sidestepping some of the pitfalls of large group collaboration.
Similarly, by creating interconnections between groups through mutual membership,
interlocks influence the overall structural properties of the network.
In turn, those structural properties influence the speed and flow of information diffusion through the entire population of collaborators.

@@

\subsection{Research Questions}
Our primary focus is whether interlocking network structure improves the quality of social-learning-based collaboration:
\begin{quote}
{\bfseries RQ 1.} Does interlocking structure influence solution quality in large-scale social learning?
\end{quote}

@@

The above examples all share the same social network structure of small groups connected through mutual membership: {\em interlocks}.
Interlocks have been well-studied in several other contexts.
Perhaps the most well-known is the study of intelocking directorates,
the practice of one individual sitting on the board of directors for
multiple corporations,
allowing an implicit flow of information and influence
\cite{sapinski_interlocking_2018}.
Interlocks arise in other fields as well,
including revolutionary cells
\cite{fishcer_paul_1994, healey_using_2013},
the UK suffragette movement \cite{crossley_covert_2012},
and the global network of cities \cite{taylor_world_2015}.

@@

\subsection{When Social Learning Goes Wrong}
While social learning can allow members of a group to combine their knowledge and resources,
it can also introduce new problems.
When information is repeated without sufficient independent checking,
a false belief starting with one person can be repeated and amplified until it is accepted by a large number of people:
an {\em information cascade} \cite{banerjee_simple_1992}.
Similarly, in the {\em majority illusion},
an individual may incorrectly infer that the majority of a group holds a particular belief if they are only exposed to others who do hold that belief \cite{lerman_majority_2015}.
Finally,
rather than reaching a single consensus,
social learning can cause {\polarization},
in which a group fragments into two or more factions which hold similar beliefs within each group,
but conflicting beliefs between groups
\cite{schkade_what_2007}.
In groups that successfully implement large-scale participatory decision-making,
these pitfalls must be avoided.
One of this paper's main goals is to identify potential structural or behavioral mechanisms that might allow a group to do so.

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-netdelib-results-net-boxplot.png}
\caption{TODO}
\label{fig:quality}
\begin{description}
\end{description}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-conform-transition-basin.png}
\caption{TODO}
\label{fig:transitions}
\begin{description}
\end{description}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-netdelib-converge-net-boxplot.png}
\caption{TODO}
\label{fig:convergence}
\begin{description}
\end{description}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-conform-basin.png}
\caption{TODO}
\label{fig:basins}
\begin{description}
\end{description}
\end{center}
\end{figure}

>>>>

%\author{Herminio Bodon}
%\orcid{0009-0004-3270-3621}
%\email{HerminioBodon2020@u.northwestern.edu}
%\affiliation{%
%  \institution{Northwestern University}
%  \city{Evanston}
%  \state{Illinois}
%  \country{USA}
%}
%\affiliation{%
%  \institution{University of Michigan}
%  \city{Ann Arbor}
%  \state{Michigan}
%  \country{USA}
%}

@@

In this paper, we begin with a review of the literature on large-scale participatory decision-making and related topics.
We focus, in particular, on empirical studies of successful large-scale participatory projects, looking for commonalities.
We then focus on one such commonality: social network structure,
specifically a type of network known as an interlock network.
We also narrow our focus to social learning:
the process by which individuals revise their beliefs and opinions
based on observations and interactions with others---a core part of participatory decision-making.
Finally, we develop an agent-based model
to test the effects of interlock networks on social learning
in large groups.
Our findings suggest that interlock network structure may contribute to the success of large-scale participatory projects by providing protection from information cascades and polarization.

@@

\subsection{Collaboration and Network Structure}
As we will elaborate below,
we find that real-world examples of successful large-scale participatory collaborations
often share a common trait:
interlocking social network structure.
Here, we use ``network'' in the sense of a set of entities with dyadic relationships between them \cite{newman_networks_2018}, e.g.,
individuals linked to other individuals through direct communication.
Similarly, we use ``structure'' to refer to the configuration of individual dyadic relationships and connections, rather than in the systemic or institutional sense. A roundtable discussion for example,
is structured to allow communication links between any pair of indidividuals, while a lecture has links from the lecturer to the audience but not between audience members (at least if they are polite).
An {\em interlocking network} or {\em interlock}
is a network in which groups of entities are connected (interlocked)
through shared individual constituents \cite{taylor_world_2015}.
As a much-studied example,
an interlocking directorate is a network of corporations
in which two corporations (the groups) are considered connected if any individual director (the constituent) belongs to both of their boards 
\cite{mizruchi_what_1996}.
Or as a more modern example,
collaborative interactions on Wikipedia take place on individual articles, their per-article talk pages, and on topical WikiProject pages \cite{platt_network_2018}.
In this example, articles/talk pages/projects are the groups, and individual editors are the constituents.

@@

There is considerable evidence that group composition
\cite{hill_group_1982,
nishi_inequality_2015,
robert_jr_differences_2018,
sydow_diversity_2017,
lerner_diverse_2018,
arazy_information_2011}
and social network structure
\cite{
kearns_experiments_2012,
barkoczi_social_2016,
salehi_hive:_2018,
mason_propagation_2008,
mason_collaborative_2012}
can influence the success of a collaboration.
Interlocking structure has implications for both of these factors.
By dividing an entire collaborative project into smaller groups,
interlocks reduce group size and alter group composition,
potentially sidestepping some of the pitfalls of large group collaboration.
For example, a participant who monoplizes the conversation might derail their small group, but not the entire project.
Or, a member of an underrepresented demographic would have more chances to speak as a participant in a small group of five versus a large group of one hundred.
Additionally, by creating interconnections between groups through mutual membership,
interlocks influence the overall structural properties of the network.
In turn, those structural properties influence the speed and flow of information diffusion through the entire population of collaborators.

@@

\subsection{Research Questions}
Our primary focus is whether interlocking network structure is beneficial for social-learning-based collaboration:
\begin{quote}
{\bfseries RQ 1.} Does interlocking structure influence solution quality in large-scale social learning?
\end{quote}

@@

The above examples all share the same social network structure of small groups connected through mutual membership: {\em interlocks}.
Interlocks have been well-studied in several other contexts.
Perhaps the most well-known is the study of intelocking directorates,
the practice of one individual sitting on the board of directors for
multiple corporations,
allowing an implicit flow of information and influence
\cite{sapinski_interlocking_2018}.
Interlocks arise in other fields as well,
including revolutionary cells
\cite{fischer_paul_1994, healy_using_2013},
the UK suffragette movement \cite{crossley_covert_2012},
and the global network of cities \cite{taylor_world_2015}.

@@

\subsection{When Social Learning Goes Wrong}
While social learning can allow members of a group to combine their knowledge and resources,
it can also introduce new problems.
When information is repeated without sufficient independent checking,
a false belief starting with one person can be repeated and amplified until it is accepted by a large number of people:
an {\em information cascade} \cite{banerjee_simple_1992}.
Similarly, in the {\em majority illusion},
an individual may incorrectly infer that the majority of a group holds a particular belief if they are only exposed to others who do hold that belief \cite{lerman_majority_2015}.
Finally,
rather than reaching a single consensus,
social learning can cause {\em polarization},
in which a group fragments into two or more factions which hold similar beliefs within each group,
but conflicting beliefs between groups
\cite{schkade_what_2007}.
In groups that successfully implement large-scale participatory decision-making,
these pitfalls must be avoided.
One of this paper's main goals is to identify potential structural or behavioral mechanisms that might allow a group to do so.

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-netdelib-results-net-boxplot.png}
\caption{Solution quality for 1000 runs of the agent-based model. For each run, quality is avereaged over all 100 agents.}
\label{fig:quality}
\Description[Two box plots]{The left plot shows results for the best-neighbor strategy, similar across all three networks, ranging from 0.2 to 1.0 with medians around 0.7. The right plot shows results for the conform strategy, ranging from 0.2 to 1.0. Small-world and preferential attachment have medians around 0.45, while random-pod has median about 0.8.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-conform-transition-basin.png}
\caption{Mean number of agents adopting solutions from a different basin of attraction.}
\label{fig:transitions}
\Description[Three time-series plots, one for each network]{All plots start around 30, curve downward at different rates described in the text, and reach 0 around timestep 30.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-netdelib-converge-net-boxplot.png}
\caption{Distribution of convergence time for 1000 runs of the simulation.}
\label{fig:convergence}
\Description[Two box plots]{The left, for best-neighbor ranges from about 5.0 to 22.5, with more
details described in the text. The right, for conform, ranges between 0 and 300,
with more details described in the text.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=5in]{fig-conform-basin.png}
\caption{Mean number of occupied basins of attraction over 1000 simulation runs.}
\label{fig:basins}
\Description[A time series plot]{Time ranges from time 0 to 30. Curves are shown for all three networks. Occupied basins starts at 100 and curves downward over time, with more details described in the text.}
\end{center}
\end{figure}
